---
title: "First Run Results-12-03-2025"
author: "Emily Tyszka"
date: "2025-12-03"
output: html_document
---

```{r, message=FALSE}

library(targets)
library(tidyverse)
library(ggbump)
library(kableExtra)

### Score computation
tar_load("Kat15k_alloscores")
tar_load("score_data")
tar_load("truth_data")
tar_load("forecast_data")
tar_load("percap")


Kat15k_alloscores <- bind_rows(Kat15k_alloscores, percap |> filter(K == 15000))

alloscore_data_summaries <- Kat15k_alloscores |>
  group_by(model, reference_date) |>
  summarize(alloscore = sum(components), yobs = sum(y)) |>
  group_by(reference_date) |>
  mutate(nmodels = n(),
         allo_rank = rank(alloscore, ties.method = "min"),
         allo_rank_std = 1- (allo_rank-1)/(nmodels-1)) |>
  ungroup() |>
  complete(model, reference_date) |>
  mutate(target_end_date = as.Date(reference_date) + 14)

alloscore_models <- unique(alloscore_data_summaries$model)
score_data_summaries <- score_data |>
  mutate(reference_date = target_end_date - 14) |>
  filter(model %in% alloscore_models) |>
  group_by(model, reference_date) |>
  summarize(mwis = mean(wis)) |>
  ungroup() |>
  group_by(reference_date) |>
  mutate(nmodels = n(),
         mwis_rank = rank(mwis, ties.method = "min"),
         mwis_rank_std = 1- (mwis_rank-1)/(nmodels-1)) |>
  ungroup() |>
  complete(model, reference_date) |>
  mutate(target_end_date = as.Date(reference_date) + 14)

all_scores <- left_join(alloscore_data_summaries, score_data_summaries,
                      by=c("model", "target_end_date"))

# Normalize Palette Aethetics
model_palette <- setNames(
  hcl(h = seq(15, 375, length = 18), l = 65, c = 100),
  c(unique(forecast_data$model), "per-capita"))

model_palette["COVIDhub-4_week_ensemble"] <- "thistle"
model_palette["FAIR-NRAR	"] <- "wheat"

model_linewidths <- setNames(rep(.5,
  length.out = length(model_palette)
), names(model_palette))
model_linewidths["per-capita"] <- 1.5

```


# Major changes:

-   Forecast locations changed from states+national to California counties+statewide.
-   Date range of interest:
-   Population data now sourced from the CA Dept. of Finance: <https://dof.ca.gov/forecasting/demographics/estimates/> <- tidycensus is better


# Meeting with Nick
- Big Rmd document: rename; 
      - linear; storytelling; explain decisions
      - dig into topics. explain data you're using in each section
      - include basic background data: time series, etc.
      

# One Week

```{r, message=FALSE}
tar_load(alloscore_2021.02.06)
slim_alloscore_2021.02.06 <- alloscore::slim(alloscore_2021.02.06, id_cols = c("model", "reference_date")) %>%
  bind_rows(percap %>% filter(reference_date == "2021-12-06"))

## for visual context
## curve(dnorm(x, mean=15000, sd=3000), from=0, to=30000)
ks <- seq(5000, 250000, by=2000)
kprobs <- data.frame(K = ks,
                     weight = dnorm(ks, mean=150000, sd=30000)/sum(dnorm(ks, mean=150000, sd=30000)))


ias <- slim_alloscore_2021.02.06 |>
  group_by(model, K) |>
  summarize(`Allocation score` = sum(components)) |>
  left_join(kprobs) |>
  group_by(model) |>
  summarize(IAS_wt = sum(`Allocation score`*weight, na.rm=TRUE),
            IAS = mean(`Allocation score`)) |>
  arrange(IAS)

as <- slim_alloscore_2021.02.06 |>
  filter(K == 150000) |>
  group_by(model) |>
  summarize(AS = sum(components)) |>
  arrange(AS)

```

```{r, message=FALSE}
wis_vals <- all_scores |>
  filter(target_end_date == "2021-02-06", !is.na(mwis)) |>
  select(model,
# can uncomment the two lines below to check they match
#         AS = alloscore,
#         `AS rank` = allo_rank,
         MWIS = mwis)

left_join(ias, as) |>
  left_join(wis_vals, by="model") |>
  rename(Model = model,
         `IAS centered at 15k` = IAS_wt,
         `IAS uniform` = IAS) |>
  select(Model, AS, MWIS, `IAS centered at 15k`, `IAS uniform`) |>
  #mutate(Model = gsub("_", "\\\\_", Model)) |>
  #mutate(Model = sprintf("\\texttt{%s}", Model)) |>
  arrange(AS) |>
  kableExtra::kbl(caption = "For one illustrative week, a comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS), All metrics are shown for 10 models that made forecasts of hospital admissions for 2021-02-06. Results are sorted by AS. For all metrics, lower scores indicate better accuracy.",
                 digits = 0) |> kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r}
all_scores |>
  filter(target_end_date == "2021-02-27", !is.na(mwis)) |>
  select(model,
         AS = alloscore,
         `AS rank` = allo_rank,
         MWIS = mwis,
         `MWIS rank` = mwis_rank) |>
  arrange(`AS rank`) |>
  kableExtra::kbl(label = "tab:one-week-results",
                 caption = "Allocation scores (AS) and mean weighted interval scores (MWIS) for each model' as forecasted's forecasts for ") |> kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r}
source("R/run-alloscore.R")
source("R/plot_functions.R")
locs_to_plot <- truth_data |>
  dplyr::filter(target_end_date == "2021-02-20", full_location_name != "California") |>
  dplyr::arrange(desc(value)) |>
  head(10) |>
  dplyr::pull(full_location_name)

tar_load(percap)
percap2 <- percap %>% filter(reference_date == "2021-02-20") %>% filter(K==150000)
tar_load(alloscore_2021.02.20)
slim_alloscore_2021.02.20 <- alloscore::slim(alloscore_2021.02.20, id_cols = c("model", "reference_date")) %>%
  bind_rows(percap %>% filter(reference_date == "2021-02-20") %>% filter(K =="150000"))

plot_hosp(models = c("COVIDhub-4_week_ensemble", "FAIR-NRAR	"),
          start_date = "2021-02-06",
          stop_date = "2021-03-27",
          f_date = "2021-02-20",
          te_date = "2022-02-20",
          ## highest locations in terms of hospitalizations
          locations = locs_to_plot,
          f_width1 = 4, f_alpha = .4,
          facet_ncol=2,
          allocations = FALSE,
         forecasts_hosp = forecast_data %>% filter(reference_date == "2021-02-20") %>%
             dplyr::mutate(
               abbreviation = factor(full_location_name, levels=locs_to_plot, ordered=TRUE),
               geo_value = abbreviation,
               alloscores = slim_alloscore_2021.02.20) %>%
             dplyr::mutate(
               abbreviation = factor(full_location_name, levels=locs_to_plot, ordered=TRUE)),
          truth = truth_data %>%
            dplyr::mutate(
              abbreviation = factor(full_location_name, levels=locs_to_plot, ordered=TRUE),
              geo_value = abbreviation),
          f_colors = model_palette[c("COVIDhub-4_week_ensemble", "FAIR-NRAR	")],
          one_K = 150000) 
          + labs(x = NULL)

forecasts_hosp <- forecast_data %>% filter(reference_date == "2021-02-20") %>%
             dplyr::mutate(
               abbreviation = factor(full_location_name, levels=locs_to_plot, ordered=TRUE),
               geo_value = abbreviation),
               alloscores = slim_alloscore_2021.02.20) %>%
             dplyr::mutate(
               abbreviation = factor(full_location_name, levels=locs_to_plot, ordered=TRUE))

locs_to_plot <- truth_data |>
  dplyr::filter(target_end_date == "2022-02-27", location != "06") |>
  dplyr::arrange(desc(value)) |>
  head(10) |>
  dplyr::pull(location_name)
```


